# pre-process pharmacogenomics data {crxg_prepro}
## center and scale data {crxg_cs}
Next, I center and scale the log-transformed ic50s to compare distributions.

```{r}
crxg <- crxg %>%
  group_by(drug_id) %>%
  mutate(rob_z = (ln_ic50 - median(ln_ic50))/mad(ln_ic50),
         rob_center = (ln_ic50 - median(ln_ic50)))

crxg %>% 
  ggplot(aes(rob_z, group = drug_id)) + 
  geom_density(alpha = 1, size = 0.2) + 
  theme_classic() + 
  labs(x = "centered and scaled ln(IC50)") + 
  ggsave("norm_ic50.pdf", width = 3, height = 2)
```

For reference, I also plot the centered (unscaled) ln(IC50) values.

```{r}
crxg %>% 
  ggplot(aes(rob_center, group = drug_id)) + 
  geom_density(alpha = 1, size = 0.2) + 
  theme_classic() + 
  labs(x = "centered ln(IC50)") + 
  ggsave("centered_ic50.pdf", width = 3, height = 2)
```

## filter underutilized compounds {crxg_qc_compounds}

I wonder if there are compounds which were only seldom screened. I remove these drugs and work with the residual dataset.

```{r}
cutoff <- 850

keep_drugs <-crxg %>% 
  ungroup %>% filter(!is.na(drug_id), !is.na(cosmic_id)) %>%
  select(cosmic_id, drug_id) %>% distinct() %>% group_by(drug_id) %>% summarise(n = n()) %>% 
  filter(n > cutoff) %>%  
  drop_na() %>% .$drug_id

crxg %>% 
  ungroup %>% filter(!is.na(drug_id), !is.na(cosmic_id)) %>%
  select(cosmic_id, drug_id) %>% distinct() %>% group_by(drug_id) %>% summarise(n = n()) %>% 
  arrange(n) %>%
  mutate(drug_id = factor(drug_id, levels =  .$drug_id)) %>% 
  ggplot(aes(drug_id, n)) + 
  geom_point() + 
  theme_classic() + 
  #scale_y_log10() + 
  theme(axis.text.x = element_blank()) + 
  geom_hline(yintercept = cutoff) + 
  ggsave("filter_drugs.pdf", width = 3, height = 2)
```

I overwrite the original object after QC. 

```{r}
crxg <- crxg %>% 
  ungroup %>% filter(!is.na(drug_id), !is.na(cosmic_id), drug_id %in% keep_drugs)
```


Finally, I save my result for further analysis

```{r}
save(crxg, file = "crxg.Rdata")
```


