---
title: "1_tidy_cancerrxgene_data"
author: "Niklas Rindtorff"
date: "3/5/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction 
#Load packages

```{r}
library(readxl)
library(tidyverse)
library(stringr)
library(janitor)
library(here)
library(googlesheets)
library(ggridges)
```

#Import drug vulnerability data

```{r}

crxg <- read_excel(here("local_data/v17.3_fitted_dose_response.xlsx")) %>% dplyr::select(-c(1,2)) %>% `colnames<-`(colnames(.) %>% tolower()) %>%
  full_join(., read_excel(here("local_data/Screened_Compounds.xlsx")) %>% janitor::clean_names()) %>%
  full_join(., read_excel(here("local_data/Cell_Lines_Details.xlsx"), sheet = "Cell line details") %>% `colnames<-`(colnames(.) %>% tolower() %>% str_replace_all(., " ", "_"))  %>%  dplyr::select(c(1,2,8,9,10,11,12,13)) %>% `colnames<-`(c("sample_name", "cosmic_id", "gdsc_tissue_1", "gdsc_tissue_2", "cancer_type", "msi", "medium", "growth_properties")))
  
```

I preprocess the drug names.

```{r, message=FALSE, warning=FALSE}
# match_drugs <- read_csv(here("data/nci_match_compounds.csv"), col_names = FALSE) %>% 
#   dplyr::select(drug_name = X1) %>% 
#   mutate(drug_name = drug_name %>% tolower())
# 
# mda_drugs <- read_csv(here("data/md_anderson_compounds.csv"), col_names = TRUE) %>% 
#   filter(include_1 == TRUE) %>% 
#   mutate(drug_name = drug %>% tolower()) %>% 
#   dplyr::select(drug_name)
# 
# default_drugs <- c(#"paclitaxel",
#                    "cisplatin")

crxg <- crxg %>% 
  mutate(drug_name = drug_name %>% tolower()) #%>% 
 # filter(drug_name %in% match_drugs$drug_name | drug_name %in% mda_drugs$drug_name | drug_name %in% default_drugs)
  
```


#Scale data

Now I plot the distribution of ln_ic50s for all drugs. They are all over the place.

```{r}
crxg %>% 
  ggplot(aes(ln_ic50, group = drug_id)) + 
  geom_density(alpha = 1, size = 0.2) + 
  theme_classic() + 
  labs(x = "raw ln(IC50)") + 
  ggsave("raw_ic50.pdf", width = 3, height = 2)
```

Now I scale the ic50s to compare distributions

```{r}
crxg <- crxg %>%
  group_by(drug_id) %>%
  mutate(rob_z = (ln_ic50 - median(ln_ic50))/mad(ln_ic50),
         rob_center = (ln_ic50 - median(ln_ic50)),
         med_ln_ic50 = median(ln_ic50, na.rm = TRUE),
         norm_ln_ic50 = ln_ic50 - med_ln_ic50)

crxg %>% 
  ggplot(aes(norm_ln_ic50, group = drug_id)) + 
  geom_density(alpha = 1, size = 0.2) + 
  theme_classic() + 
  labs(x = "centered ln(IC50)") + 
  ggsave("center_ic50.png", width = 3, height = 3)
```


#QC

I wonder if there are compounds which were only seldom screened. I remove these drugs and work with the residual dataset.

```{r}
cutoff <- 850

keep_drugs <-crxg %>% 
  ungroup %>% filter(!is.na(drug_id), !is.na(cosmic_id)) %>%
  select(cosmic_id, drug_id) %>% distinct() %>% group_by(drug_id) %>% summarise(n = n()) %>% 
  filter(n > cutoff) %>%  
  drop_na() %>% .$drug_id

crxg %>% 
  ungroup %>% filter(!is.na(drug_id), !is.na(cosmic_id)) %>%
  select(cosmic_id, drug_id) %>% distinct() %>% group_by(drug_id) %>% summarise(n = n()) %>% 
  arrange(n) %>%
  mutate(drug_id = factor(drug_id, levels =  .$drug_id)) %>% 
  ggplot(aes(drug_id, n)) + 
  geom_point() + 
  theme_classic() + 
  #scale_y_log10() + 
  theme(axis.text.x = element_blank()) + 
  geom_hline(yintercept = cutoff) + 
  ggsave("filter_drugs.pdf", width = 3, height = 2)
```

I overwrite the original object after QC. 

```{r}
crxg <- crxg %>% 
  ungroup %>% filter(!is.na(drug_id), !is.na(cosmic_id), drug_id %in% keep_drugs)
```

I extract covariates for the cell lines and export them

```{r}
covariates <- crxg %>% 
  dplyr::select(cosmic_id, gdsc_tissue_2, msi, medium, growth_properties) %>% 
  distinct() %>%
  mutate(medium = factor(medium) %>% as.numeric(),
         medium = medium -1,
         msi = factor(msi) %>% as.numeric(),
         msi = msi-1, 
         growth_properties = factor(growth_properties),
         gdsc_tissue_2 = factor(gdsc_tissue_2)) %>% 
  mutate(value = 1) %>%
  spread(gdsc_tissue_2, value, fill = 0) %>% 
  mutate(value = 1) %>%
  spread(growth_properties, value, fill = 0) %>% 
  janitor::clean_names() %>% 
  dplyr::select(-na) %>% 
  replace_na(list(msi = 1))
  
write_csv(covariates, here("data/covariates.csv")) 
```

```{r}
classes <- covariates %>% 
  dplyr::select(-cosmic_id) %>%
  colSums()
  
classes/nrow(covariates) %>% 
  sort()
```


Finally, I save my result for further analysis

```{r}
save(crxg, file = here("data/crxg.Rdata"))
```


