---
title: "pan_cancer_analysis"
author: "Niklas Rindtorff"
date: "5/13/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache=TRUE)
```

#Introduction 
#Load packages

```{r}
library(tidyverse)
library(stringr)
library(here)
library(viridis)
```

#Load data
```{r}
load(here("data/crxg.Rdata"))
```


#Correlation pan-cancer

I show the null-distribution of drug correlations for all drugs and cell lines available.

```{r}
crxgr_pan <- crxg %>% dplyr::select(drug_id, cosmic_id, rob_z) %>%
  filter(!is.na(drug_id), !is.na(cosmic_id)) %>%
  spread(drug_id, rob_z) %>% 
  dplyr::select(-cosmic_id)%>%
  corrr::correlate(method = "spearman", use = "pairwise.complete.obs") %>%
  corrr::stretch()

crxgr_pan %>%
  drop_na() %>%
  ggplot(aes(r)) + 
  geom_density() + 
  theme_classic()
```

```{r}
crxg_map <- crxg %>% dplyr::select(drug_id, cosmic_id, ln_ic50) %>%
  filter(!is.na(drug_id), !is.na(cosmic_id)) %>% 
  group_by(drug_id) %>%
  mutate(med_ln_ic50 = median(ln_ic50, na.rm = TRUE)) %>%
  mutate(norm_ln_ic50 = ln_ic50 - med_ln_ic50) %>%
  ungroup() %>%
  dplyr::select(drug_id, cosmic_id, norm_ln_ic50) %>%
  spread(drug_id, norm_ln_ic50) %>% 
  column_to_rownames("cosmic_id") %>% 
  #drop_na() %>%
  as.data.frame() 
```

# Removing duplicated drugs

It turns out that the drug *Afatinib* and *Olaparib* were screened twice. We have to pick one of the datasets to continue.

## Afatinib

```{r}
crxg_map %>% dplyr::select(low_dose = `1032`, high_dose = `1377`) %>% 
  drop_na() %>%
  ggplot(aes(low_dose, high_dose)) + 
  geom_point(alpha = 0.3) + 
  theme_classic() + 
  geom_abline(slope = 1) + 
  labs(title = "Afatinib") +
  ggsave("afatinib_correlation.pdf", width = 3, height = 3)

conc_range_afatinib <- crxg %>% 
  filter(drug_id %in% c(1032, 1377)) %>%
  dplyr::select(max_conc_micromolar, min_conc_micromolar, drug_id) %>% 
  mutate(dose  = if_else(drug_id == 1032, "low_dose", "high_dose")) %>%
  distinct() %>% 
  mutate(max_conc_micromolar = log(max_conc_micromolar),
         min_conc_micromolar = log(min_conc_micromolar))

crxg %>% 
  filter(drug_id %in% c(1032, 1377)) %>%
  mutate(dose  = if_else(drug_id == 1032, "low_dose", "high_dose")) %>%
  ggplot(aes(ln_ic50, group = drug_id, fill = dose)) + 
  geom_density(size = 0.2, alpha = 0.5) + 
  theme_classic() + 
  labs(x = "raw ln(IC50)") + 
  geom_vline(data = conc_range_afatinib, 
             aes(xintercept=max_conc_micromolar, color=dose),
             size = 0.5) +
  geom_vline(data = conc_range_afatinib, 
             aes(xintercept=min_conc_micromolar, color=dose),
             linetype="dashed",
             size = 0.5) + 
  ggsave("afatinib_lnic50_range.pdf", width = 5, height = 3)

```

I reformat the column names after removing the low-dose *Afatinib* variable. The dose ic50 estimates are shifted towards higher values in both batches. The effect size and deviation correlates well. I end up picking the replicate with the higher number of observations.

```{r}
crxg_map <- crxg_map %>% dplyr::select(-`1032`)
```

## Olaparib 

```{r}
crxg_map %>% dplyr::select(low_dose = `1017`, high_dose = `1495`) %>% 
  drop_na() %>%
  ggplot(aes(low_dose, high_dose)) + 
  geom_point(alpha = 0.3) + 
  theme_classic() + 
  geom_abline(slope = 1) + 
  labs(title = "Olapraib") +
  ggsave("olaparib_correlation.pdf", width = 3, height = 3)

conc_range_olaparib <- crxg %>% 
  filter(drug_id %in% c(1017, 1495)) %>%
  dplyr::select(max_conc_micromolar, min_conc_micromolar, drug_id) %>% 
  mutate(dose  = if_else(drug_id == 1017, "low_dose", "high_dose")) %>%
  distinct() %>% 
  mutate(max_conc_micromolar = log(max_conc_micromolar),
         min_conc_micromolar = log(min_conc_micromolar))

crxg %>% 
  filter(drug_id %in% c(1017, 1495)) %>%
  mutate(dose  = if_else(drug_id == 1017, "low_dose", "high_dose")) %>%
  ggplot(aes(ln_ic50, group = drug_id, fill = dose)) + 
  geom_density(size = 0.2, alpha = 0.5) + 
  theme_classic() + 
  labs(x = "raw ln(IC50)") + 
  geom_vline(data = conc_range_afatinib, 
             aes(xintercept=max_conc_micromolar, color=dose),
             size = 0.5) +
  geom_vline(data = conc_range_afatinib, 
             aes(xintercept=min_conc_micromolar, color=dose),
             linetype="dashed",
             size = 0.5) + 
  ggsave("olaparib_lnic50_range.pdf", width = 5, height = 3)

```

Based on this data, I decide to stick to the drug_id that covers a greater portion of the drug's ln(IC50) values. I remove the other drug_id.

```{r}
crxg_map <- crxg_map %>% dplyr::select(-`1017`)

# Mapping interpretable names after removing duplicate row afatinib
colnames(crxg_map) <- crxg %>% dplyr::select(drug_id, drug_name) %>% 
  filter(drug_id %in% colnames(crxg_map)) %>% 
  distinct() %>% .$drug_name

save(crxg_map, file = here("data/crxg_map.Rdata"))
```

# Visualize distribution of response values.

```{r}
crxg_map %>% 
  gather(drug_name, norm_ln_ic50) %>% 
  drop_na() %>% 
  mutate(drug_name = factor(drug_name)) %>%
  ggplot(aes(y = drug_name, x = norm_ln_ic50, fill = ..x..)) + 
  geom_density_ridges_gradient() + 
  scale_fill_viridis(name = "Centered ln(IC50)", option = "C") + 
  theme_ridges() + 
  labs(x = "",
       y = "",
       title = "Distribution of drug responses") + 
  ggsave("drug_response_dist.pdf", width = 5, height = 3)
```


# Missing values

I plot missing values

```{r}
crxg_map %>%
  naniar::vis_miss()
```


Without dropping NAs the dimensions are: 

```{r}
crxg_map %>% dim()
```

With dropping NAs the dimensions are: 

```{r}
crxg_map %>%
  drop_na() %>% dim()
```

I impute missing data by setting missing values to the median response value. In our case the median response value is defined to be 0. 


```{r}
crxg_map_imputed <- crxg_map %>% 
  naniar::impute_median_all() %>% 
  as.data.frame() %>% 
  magrittr::set_rownames(rownames(crxg_map)) %>% 
  rownames_to_column("cosmic_id")

save(crxg_map_imputed, file = here("data/crxg_map_imputed.Rdata"))
write_csv(crxg_map_imputed, path = here("data/drug_response.csv"))
```


Now, I plot the data. 

```{r}
types_of_interest <- (crxg %>% 
    drop_na() %>%
    dplyr::select(cosmic_id, cancer_type ) %>% 
    distinct() %>% count(cancer_type) %>% arrange(desc(n)) %>% filter(n > 30) %>% .$cancer_type)


anno_col  = crxg %>% 
  drop_na() %>%
                  dplyr::select(cosmic_id, cancer_type ) %>% 
                  distinct() %>%
  mutate(cancer_type = if_else(cancer_type %in% types_of_interest,
    cancer_type, 
    "Other")) %>%
  as.data.frame() %>%
  remove_rownames() %>%
 # dplyr::select(-cancer_type) %>%
                  column_to_rownames("cosmic_id")

anno_row  = crxg %>% 
  drop_na() %>%
                  dplyr::select(drug_name, target_pathway) %>% 
                  distinct() %>%
  as.data.frame() %>%
  remove_rownames() %>%
                  column_to_rownames("drug_name")

# defining colors - always fun
cancer_type_color <- RColorBrewer::brewer.pal(length(types_of_interest)+1, "Paired")
names(cancer_type_color) <- c(types_of_interest, "Other")

target_pathway <- RColorBrewer::brewer.pal(length(anno_row$target_pathway %>% unique()), "Set1")
names(target_pathway) <- anno_row$target_pathway %>% unique()

anno_color <- list(cancer_type = cancer_type_color,
                   target_pathway = target_pathway)

crxg_map_imputed %>%
  dplyr::select(-cosmic_id) %>%
  as.data.frame() %>%
  drop_na() %>%
  t() %>%
  pheatmap::pheatmap(na_col = "grey", scale = "column",
                     annotation_col = anno_col, 
                     annotation_row = anno_row,
                     annotation_colors = anno_color,
                     annotation_names_row = FALSE,
                     show_rownames = TRUE, 
                     show_colnames = FALSE,
                     cutree_rows = 5, 
                     filename = "crxg_no_allocation.pdf", 
                     width = 9, height = 4.8
                     )
                     
```




