---
title: "subtype_specific_anaysis"
author: "Niklas Rindtorff"
date: "5/13/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache=TRUE)
```

#Introduction 

I want to simulate the current inclusion criteria into NCI Match. 

#Load packages

```{r}
library(tidyverse)
library(stringr)
library(here)
```

#Import data

```{r}
load(here("data/crxg.Rdata"))
load(here("data/crxg_map.Rdata"))
```


#Filtering cell lines
The current result only return the minimal r for a pan-cancer approach. However, I am interested in anti-correlation in clinically relevant subgroups. 

First I see how many cell lines are present for every subgroup. I kick out all cancertypes with less than 10 cell lines per entity and no clear annotation.

```{r}
keep_types <- crxg %>% ungroup %>% select(cancer_type, cosmic_id) %>% distinct() %>% group_by(cancer_type) %>% 
  summarise(n = n()) %>% 
  arrange(n) %>% filter(n > 10) %>% 
  filter(!grepl(cancer_type, pattern = "UNABLE")) %>% drop_na() %>% .$cancer_type 

crxg %>% ungroup %>% filter(!is.na(drug_id), !is.na(cosmic_id)) %>%
  ##modification to group by gdsc_tissue_1 annotation 
  #dplyr::select(-cancer_type) %>% 
  #rename(cancer_type = gdsc_tissue_1) %>% 
  ##
  select(cancer_type, cosmic_id) %>% distinct() %>% group_by(cancer_type) %>% summarise(n = n()) %>% 
  arrange(n) %>% 
  mutate(cancer_type = factor(cancer_type, levels = .$cancer_type)) %>%
  ggplot(aes(cancer_type, n)) + 
  geom_point() + 
  theme_classic() + 
  scale_y_log10() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  geom_hline(yintercept = 10) + 
  xlab("Cance Type") + 
  ggsave("filter_cell_lines.pdf", width = 3, height = 2)
```

I remove all cancer entities that do not make it past the threshold.

```{r}
crxgf <- crxg %>% filter(cancer_type %in% keep_types)
save(crxgf, file = "crxgf.Rdata")
```

#Correlate responses for each cancer type

I define a function to calculate spearman correlations for multiple subgroups.
Then I plot the null distribution of all correlation coefficients for every tissue type.

```{r}
calc_subgroup_cor <- function(df){
tmp <- df %>% dplyr::select(drug_id, cosmic_id, rob_z) %>%
  spread(drug_id, rob_z) %>% 
  dplyr::select(-cosmic_id)%>%
  corrr::correlate(method = "spearman", use = "pairwise.complete.obs") %>%
  corrr::stretch() %>% 
  unite(pair, x, y)

return(tmp)
}

crxgr_type <-crxgf %>% 
  filter(!is.na(drug_id), !is.na(cosmic_id), cancer_type %in% keep_types) %>%
  group_by(cancer_type) %>% 
  do(calc_subgroup_cor(.))

crxgr_type %>% 
  ggplot(aes(r)) + 
  geom_density(aes(group = cancer_type)) + 
  theme_classic() + 
  xlab("Spearman correlation coefficient") + 
  ggsave("null_r_type.pdf", width = 3, height = 2)
```


Now I link the group-wise effects with the anti-correlation scores.

```{r, eval = FALSE}
#crxgr_type
crxgra_type <- crxgr_type %>%  
  ungroup() %>% #crxgr_type with annotation 
  #dplyr::sample_n(1000) %>%
  #distinct(., keep_all = TRUE) %>% 
  separate(pair, c("d1", "d2"), remove = TRUE) %>% #I keep only the first id for each row
  mutate(d1 = d1 %>% as.numeric) %>% 
  mutate(d2 = d2 %>% as.numeric)  %>% 
  rowwise() %>%
  mutate(pair = paste(c(d1, d2) %>% sort(), collapse =  "_")) %>%
  drop_na() %>% 
  dplyr::select(-d1, -d2) %>% 
  distinct(cancer_type, r, pair) %>% 
  separate(pair, c("d1", "d2"), remove = FALSE) %>% #Now the columns are ordered systematically
  gather(number, drug_id, -r, -cancer_type, -pair) %>%
  
  #rowwise() %>%
  #mutate(pair = str_split(pair, pattern = "_") %>% simplify() %>% as.numeric() %>% sort() %>% paste(., collapse =  "_")) %>% 
  ungroup() %>%
######## commenting out this phrase since cancer_effects are not assesed
  #left_join(., cancer_type_effects) %>%
########
  left_join(., crxg %>% dplyr::select(drug_id, drug_name, target) %>% 
              ungroup %>% mutate(drug_id = as.character(drug_id)) %>% 
              distinct())
```

I perform a quick save

```{r, eval = TRUE}
save(crxgra_type, file="crxgra_type.Rdata")
```

```{r}
load(file="crxgra_type.Rdata")
```

I run another set of formating step of the dataset

```{r}
crxgra_type <- crxgra_type %>% separate(pair, c("d1", "d2"), remove = FALSE) %>% 
  dplyr::select(-number, - drug_id, -drug_name, -target) %>% distinct() %>%
  left_join(crxg %>% dplyr::select(drug_id, drug_name, target) %>%
              mutate(d1 = drug_id %>% as.character) %>% 
              distinct() %>% dplyr::select(-drug_id)) %>% rename(d1_drug_name = drug_name,
         d1_target = target) %>% 
  left_join(crxg %>% dplyr::select(drug_id, drug_name, target) %>%
              mutate(d2 = drug_id %>% as.character) %>% 
              distinct() %>% dplyr::select(-drug_id)) %>% rename(d2_drug_name = drug_name,
         d2_target = target) %>% distinct()

crxgr_type <- crxgr_type %>% separate(pair, c("d1", "d2"), remove = FALSE) %>% 
   distinct() %>%
  left_join(crxg %>% dplyr::select(drug_id, drug_name, target) %>%
              mutate(d1 = drug_id %>% as.character) %>% 
              distinct() %>% dplyr::select(-drug_id)) %>% rename(d1_drug_name = drug_name,
         d1_target = target) %>% 
  left_join(crxg %>% dplyr::select(drug_id, drug_name, target) %>%
              mutate(d2 = drug_id %>% as.character) %>% 
              distinct() %>% dplyr::select(-drug_id)) %>% rename(d2_drug_name = drug_name,
         d2_target = target) %>% distinct()
```

#GI cancer Case Study 

```{r}
crxgr_type %>% 
  ggplot(aes(x = r, y = cancer_type)) + 
  geom_density_ridges(aes(group = cancer_type), rel_min_height = 0.005) + 
  theme_classic() + 
  xlab("Spearman correlation coefficient") + 
  theme_ridges()
```

I select a subset of GI cancer types and remove compounds show an overall low response in that specific cancer type

```{r}
# crxg %>% 
#   filter(!is.na(drug_id), !is.na(cosmic_id), cancer_type %in% c("COAD/READ", "ESCA", "PAAD", "STAD")) %>% 
#   group_by(drug_name) %>% 
#   summarise(var = var(rob_z)) %>% 
#   filter(var > 2) %>%
#   ggplot(aes(reorder(drug_name, -var), var)) + 
#   geom_point()

crxg %>% 
  filter(!is.na(drug_id), !is.na(cosmic_id)) %>% 
  dplyr::select(drug_name, cancer_type, ln_ic50) %>% 
  nest(-drug_name, -cancer_type) %>% 
  left_join(., 
    crxg %>% 
    filter(!is.na(drug_id), !is.na(cosmic_id)) %>% 
    dplyr::select(drug_name, ln_ic50) %>% 
    nest(-drug_name, .key = pan_data)) %>% 
  filter(cancer_type %in% c("COAD/READ", "ESCA", "PAAD", "STAD")) %>% 
  mutate(data = map(data, ~ as.data.frame(.x))) %>%
  mutate(pan_data = map(pan_data, ~ as.data.frame(.x))) %>%
  mutate(tmp = map2(data, pan_data, ~ ks.test(.x, .y)))
  
  
#crxg %>%
#  filter(!is.na(drug_id), !is.na(cosmic_id)) %>% 
  
  
```



```{r}
crxgf %>% 
  #filter(!is.na(drug_id), !is.na(cosmic_id), cancer_type %in% keep_types) %>%
  #filter(cancer_type == "COAD/READ") %>% 
  filter(!is.na(drug_id), !is.na(cosmic_id), cancer_type %in% c("COAD/READ", "ESCA", "PAAD", "STAD")) %>%
  dplyr::select(drug_id, cosmic_id, rob_z) %>%
    spread(drug_id, rob_z) %>% 
    dplyr::select(-cosmic_id)%>%
    cor(method = "spearman", use = "pairwise.complete.obs") %>% 
  pheatmap::pheatmap(filename = "gi_cancer_cor.pdf",
                     width = 10, height = 10)
  
  
```


I plot an example of a strongly correlating drug pair and an anti-correlating drug-pair 

```{r}
crxg %>% filter(drug_id %in% c(179, 1377), cancer_type == "COAD/READ") %>% 
  select(drug_id, ln_ic50, cosmic_id) %>% 
  spread(drug_id, ln_ic50) %>% 
  ggplot(aes(`179`, `1377`)) + 
  geom_point() + 
  theme_classic() +
  geom_smooth(method=lm, se=FALSE, color = "black") +
  ggtitle("5FU and Afatinib (EGFRi)")
```

For contrast, I plot a drug pair which correlates very strong. Both compounds have the same mechanism of action.


```{r}
#1372_1526
crxg %>% filter(drug_id %in% c(1526, 1372), cancer_type == "COAD/READ") %>% 
  select(drug_id, ln_ic50, cosmic_id) %>% 
  spread(drug_id, ln_ic50) %>% 
  ggplot(aes(`1526`, `1372`)) + 
  geom_point() + 
  theme_classic() + 
  geom_smooth(method=lm, se=FALSE, color = "black") +
  ggtitle("Trametinib and Refamitinib") 
```


#Abadoned Approaches 
##Estimate treatment effect size

Now I want to link a correlation pair with a median relative effect for each substance of the pair in its respective subset.
First I build a table that describes the drug effect for a defined subset. 
This approach is not perfect yet, as it is probably highly dependent on the size of each cancer_type group.

```{r}
cancer_type_effects <- crxg %>% group_by(cancer_type, drug_id) %>% summarise(type_effect = median(rob_z, na.rm = TRUE)) %>% mutate(drug_id = as.character(drug_id))

cancer_type_effects%>% 
  ggplot(aes(type_effect)) + 
  geom_density(aes(group = cancer_type)) + theme_classic()
```



Now I calculate a score for each drug pair 

```{r}
crxgrae_type <- crxgra_type %>% 
  ungroup() %>% 
  #sample_n(1000) %>%
  group_by(cancer_type, pair) %>%
  mutate(effect_sum = sum(type_effect)) %>% 
  ungroup() %>% 
  mutate(score = (r-1) * effect_sum) 

crxgrae_type %>% 
  #filter(cancer_type == "COAD/READ") %>%
  ggplot(aes(effect_sum, r)) + 
  #geom_point(aes(color = score), alpha = 0.2) + 
  geom_density_2d() +
  theme_classic() + 
  geom_smooth()
```

I take a look at some examples 

```{r}
coi <- c("179_1494") #, #Irinotecan + 5-FU
         #)
# 326_1377 hit
# 179_1377 5-FU and EGFR

crxgrae_type %>% 
  filter(cancer_type == "COAD/READ") %>% 
  arrange(r) %>% 
  mutate(pair = factor(pair, levels = .$pair %>% unique())) %>% 
  ggplot(aes(pair, r)) + 
  geom_point() + 
  #geom_density() + 
  theme_classic() + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```


##Draw networks of anticorrelated drug pairs


```{r}
crxgr_type %>% drop_na() %>% group_by(cancer_type) %>% 
  mutate(m = median(r, na.rm = TRUE), mad = m-3*mad(r, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(hit = if_else(r < mad, TRUE, FALSE)) %>% 
  filter(hit == TRUE) %>% 
  filter(cancer_type == "LUAD") %>% 
  dplyr::select(d1_drug_name, d2_drug_name, r) %>% 
  #Afatinib is a drug with multiple ids. To remove these duplicates I simply build an average.
  group_by(d1_drug_name, d2_drug_name) %>% 
  summarise(r = mean(r)) %>%
  spread(d1_drug_name, r) %>% 
  column_to_rownames("d2_drug_name") %>% 
  pheatmap::pheatmap(cluster_rows = F, cluster_cols = F, na_col = "white")
```

```{r}
crxgra_type %>% drop_na() %>% group_by(cancer_type) %>% 
  mutate(m = median(r, na.rm = TRUE), mad = m-3*mad(r, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(hit = if_else(r < mad, TRUE, FALSE)) %>% 
  filter(hit == TRUE) %>% 
  filter(cancer_type == "COAD/READ") %>% 
  #dplyr::select(d1_drug_name, d2_drug_name, r) %>% 
  #Afatinib is a drug with multiple ids. To remove these duplicates I simply build an average.
  group_by(d1_drug_name, d2_drug_name) %>% 
  summarise(r = mean(r)) %>% 
  as_tbl_graph(directed = FALSE) %>% 
  ggraph(layout = 'kk') + 
  geom_edge_fan( show.legend = FALSE, width = 1.5) +
  geom_node_point(repel = TRUE, alpha = 1, size = 3) + 
  geom_node_text(aes(label = name), size = 3)  +
  theme_graph() + 
  ggsave("graph_coadread.pdf")
  
```

```{r}
crxgra_type %>% drop_na() %>% group_by(cancer_type) %>% 
  mutate(m = median(r, na.rm = TRUE), mad = m-3*mad(r, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(hit = if_else(r < mad, TRUE, FALSE)) %>% 
  filter(hit == TRUE) %>% 
  filter(cancer_type == "LUAD") %>% 
  #dplyr::select(d1_drug_name, d2_drug_name, r) %>% 
  #Afatinib is a drug with multiple ids. To remove these duplicates I simply build an average.
  group_by(d1_drug_name, d2_drug_name) %>% 
  summarise(r = mean(r)) %>% 
  as_tbl_graph(directed = FALSE) %>% 
  ggraph(layout = 'kk') + 
  geom_edge_fan( show.legend = FALSE, width = 1.5) +
  geom_node_point(repel = TRUE, alpha = 1, size = 3) + 
  geom_node_text(aes(label = name), size = 3)  +
  theme_graph() + 
  ggsave("graph_luad.pdf")
```

